Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-01 EconGrowth.Rnw
Working on a plot that will show the parameters from the Cobb-Douglas curve fits (lambda, alpha, beta). Still not working yet.

<<setup, include=FALSE>>=
require(knitr)
require(lattice)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  tidy=FALSE,
  fig.width=6,
  fig.height=4,
  comment=NA)

@@

<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", echo=FALSE>>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "black") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = layoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"),
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

@@

<<CD Results, results='asis', echo=FALSE>>=
print(cobbDouglasParamsTable(), caption.placement="top", sanitize.colnames.function = identity)
@

<<CD_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Cobb-Douglas (without energy) results.", echo=FALSE>>=
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the models
  dataTable <- cbind(dataTable, predictions) #Now add the predictions column to the data. Column name = "predGDP".
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1) #line widths. 0 means no line.
  colors <- c("black", "black") #line and symbol colors
  symbols <- c(1, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"),
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "Predicted")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,8), #y axis limits
                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createCDLatticeGraph()
@

@@

<<cobbDouglasEnergy function, include=TRUE, echo=FALSE>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string to be used in table captions reprsenting the type of energy. Typically, "Q", "X", or "U"
##
cobbDouglasEnergyData <- function(countryName, energyType){

@@

\subsection{CES with $Q$}
<<CESq analysisQ>>=
countryName <- "US"
energyType <- "Q"
tableCESq <- cesTable(countryName, energyType)

@@

<<CESq Results, results='asis'>>=
print(tableCESq, caption.placement="top")

>>>>

<<setup, include=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  tidy=FALSE,
  fig.width=6,
  fig.height=4,
  comment=NA)

@@

<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", echo=FALSE, eval=FALSE>>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = layoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

cobbDouglasCountryRowsForParamsGraph <- function(countryName){
  dataCD <- cobbDouglasData(countryName)
  #Create three rows, one for each parameter.
  lambdaRow <- cbind(dataCD["- 95% CI", "lambda"], 
                     dataCD["CD", "lambda"], 
                     dataCD["+ 95% CI", "lambda"], 
                     countryName, 
                     "lambda")
  alphaRow <- cbind(dataCD["- 95% CI", "alpha"],
                    dataCD["CD", "alpha"],  
                    dataCD["+ 95% CI", "alpha"],
                    countryName,
                    "alpha")
  betaRow <- cbind(dataCD["- 95% CI", "beta"],   
                   dataCD["CD", "beta"],   
                   dataCD["+ 95% CI", "beta"],
                   countryName,
                   "beta")
  #Now combine into a table
  table <- rbind(lambdaRow, alphaRow, betaRow)
  colnames(table) <- c("- 95% CI", "value", "+ 95% CI", "country", "parameter")
  return(table)
}

@@

<<CD Params Table, results='asis', echo=FALSE, eval=FALSE>>=
print(cobbDouglasParamsTable(), caption.placement="top", sanitize.colnames.function = identity)
@

<<CD_Params_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Cobb-Douglas (without energy) model parameters.">>=
createCDParamsGraph <- function(){
  # Create the data table that we want. This table has the following columns
  # -95% CI, value, +95% CI, country abbrev, parameter (lambda, alpha, or beta)
  dataTable <- do.call("rbind", lapply(countryAbbrevs, cobbDouglasCountryRowsForParamsGraph))

  print(dataTable)

  graph <- segplot(value ~ country | parameter, centers=value, data=dataTable)  
  return(graph)

}

createCDParamsGraph()

@

<<CD_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Cobb-Douglas (without energy) results.", echo=FALSE, eval=FALSE>>=
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the models
  dataTable <- cbind(dataTable, predictions) #Now add the predictions column to the data. Column name = "predGDP".
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1) #line widths. 0 means no line.
  colors <- c("black", "black") #line and symbol colors
  symbols <- c(1, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "Predicted")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,8), #y axis limits
                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createCDLatticeGraph()
@

@@

<<cobbDouglasEnergy function, include=TRUE, echo=FALSE, eval=FALSE>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string to be used in table captions reprsenting the type of energy. Typically, "Q", "X", or "U"
##
cobbDouglasEnergyData <- function(countryName, energyType){

@@

\subsection{CES with $Q$}
<<CESq analysisQ, eval=FALSE>>=
countryName <- "US"
energyType <- "Q"
tableCESq <- cesTable(countryName, energyType)

@@

<<CESq Results, results='asis', eval=FALSE>>=
print(tableCESq, caption.placement="top")
